---
title: "Assignment3"
author: "RF"
date: "3/6/2018"
output: html_document 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Multivariate linear models

In this assignment you will investigate the relation between different aspects of IQ and symptom severity in ASD (as measured by ADOS). The assignment is meant to make you practice on linear models, multiple predictors and interactions.

The data you will use is the outcome of the psychological testing of the children you analyzed in methods 3.
Data: https://www.dropbox.com/s/hom2qnmp9hyfwmu/Assignment3Data.csv?dl=0
The data consists of ChildID, gender, age, diagnosis (ASD), symptom severity (ADOS), several aspects of IQ (NonVerbal/Perceptual,Verbal,Social) and the visit at which the tests were administered. The questions should be answered by focusing on visit 1.

```{r read data}
# port=read.csv("Assignment3Data.csv")
# visit1=port[port$Visit== 1,]
# 
# #rename variables because I keep forgetting writing capitals
# mine= data.frame(id= visit1$ChildID, visit=visit1$Visit, gender= visit1$Gender, 
#                  age_months= visit1$AgeMonths, ASD=visit1$ASD, ADOS=visit1$ADOS, 
#                  nonverbalIQ= visit1$NonVerbalIQ, verbalIQ= visit1$VerbalIQ, socialIQ=visit1$SocialIQ)
# 
# write.csv(mine, "Assignment3Data_m.csv", row.names = F) #this has only visit1

port=read.csv("Assignment3Data_m.csv")

library(rethinking)

```


The questions you will have to answer are:

# Question 1
1. Assess the relation between symptom severity and IQ (focus on visit 1 and children with ASD) and report the model, a plot of the model, a couple of lines describing the quality of the model and interpreting the results. P.S. Should you scale? P.P.S. Don't forget to motivate your priors. P.P.P.S. At least one plot for results and a plot for quality of each model (here and in the next questions) would be appreciated.
1.1. Verbal IQ and ADOS
1.2. Non Verbal IQ and ADOS
1.3. Social IQ and ADOS

```{r norm}
#using visit 1 and children with ASD
one= port[port$ASD== 1,]

#standardize (mean of 0, sd of 1)
one$nonverbalIQ_s=(one$nonverbalIQ - mean(one$nonverbalIQ)) / sd(one$nonverbalIQ)
one$verbalIQ_s= (one$verbalIQ - mean(one$verbalIQ)) / sd(one$verbalIQ)
one$socialIQ_s= (one$socialIQ - mean(one$socialIQ)) / sd(one$socialIQ)

```

model:
ADOS ~ verbalIQ

ADOS ~ Normal(mu, sigma) 
mu = alpha + beta*verbalIQ

scaled priors
alpha ~ Normal (15, 2 ) #this should be around the mean of ADOS?
beta ~ Normal(-0.5, 0.5) #this should be how much I expect ADOS to be bigger/lower if I move 1 on the IQ scale
sigma ~ Uniform(0, 50) #the sd - just keep it a weak prior


```{r verbal}


#plotting priors ☻ 
#plot prior for alpha - ADOS 
curve( dnorm( x , 15 , 2) , from=5 , to=25 ) 
#when verbalIQ is 0 (that's the mean!) I expect ADOS to be between 10 and 20 #weak prior

#prior for beta 
curve( dnorm( x , -0.5 , 0.5) , from=-2 , to=2 )
#for each increase in unit in verbal IQ I expect ADOS to decrease 0.5 # weak prior, I left opportunity for ADOS to increase (which would mean the higher the verbal IQ the "more autistic")

#prior for sigma 
curve( dnorm( x , 0 , 50) , from=-150 , to=150 )
#a weak uniform prior


plot(y=one$ADOS, x=one$verbalIQ_s)
#does this make any sense if one of the variables is standardized??


# # fit model
# model <- map(
#   alist(
#     ADOS ~ dnorm( mu , sigma ) ,
#     mu <- a + b*verbalIQ ,
#     a ~ dnorm( 15 , 2 ) ,
#     b ~ dnorm( 0 , 0.5 ) ,
#     sigma ~ dunif( 0 , 50 )
#   ) ,  data=one )
# 
# 
# #output with variance-covariance matrix
# precis( model , corr=TRUE ) #big corr (0.9) between a and b, use standardized


# fit model
model_v <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + b*verbalIQ_s ,
    a ~ dnorm( 15 , 2 ) ,
    b ~ dnorm( -0.5 , 0.5 ) ,
    sigma ~ dunif( 0 , 5 )
  ) ,  data=one )


#output with variance-covariance matrix
precis( model_v , corr=TRUE ) 


```

Visualize

```{r prior+post}
#plotting the priors against the posteriors
post = extract.samples( model_v , n=1e4 )
dist_a= data.frame(poste= post$a, prior= rnorm(1e4, 15, 2)) #distribution of alpha (posterior and prior)

ggplot(data= dist_a)+ 
  geom_line(aes(poste), stat= "density") +
  geom_line(aes(prior), stat= "density", colour= "red")+
  labs(x="data", y="density")


post = extract.samples( model_v , n=1e4 )
dist_b= data.frame(poste= post$b, prior= rnorm(1e4, 0 , 0.5)) #distribution of alpha (posterior and prior)

ggplot(data= dist_b)+ 
  geom_line(aes(poste), stat= "density") +
  geom_line(aes(prior), stat= "density", colour= "red")+
  labs(x="data", y="density")
```


```{r plot verbal}

#looking at the predictions of the model when verbal IQ is 0
post <- extract.samples( model_v , n=2000 )
at_zero= post$a + post$b * 0
dens( at_zero , col=rangi2 , lwd=2)
HPDI( at_zero , prob=0.89 )
  
#x axis (length.out: how many points)
verbalIQ_seq <- seq( from= min(one$verbalIQ_s) , to=max(one$verbalIQ_s) , length.out=1000 ) 
pred=data.frame(verbalIQ_s=verbalIQ_seq)

mu <- link( model_v , n=1e4, data= pred)
#link: distribution of mu for each unique value on the x axis (defined in verbalIQ_seq)
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 ) #small shaded area, uncertainty of mu
sim_ADOS <- sim( model_v , data=pred ) #creating new outcome values based on the model and using my sequence
ADOS_PI <- apply( sim_ADOS , 2 , PI , prob=0.89 ) #big shaded area, uncertainty of mu + sigma 

#run all the lines together
plot( ADOS ~ verbalIQ_s , one , col=col.alpha(rangi2,0.5) ) #original data
lines( verbalIQ_seq , mu.mean ) #most likely regression line (~lm)
shade( mu.PI , verbalIQ_seq ) #small shaded area (89% of the regression lines is here)
shade( ADOS_PI , verbalIQ_seq ) #big shaded area (89% of predicted values)



```

model:
ADOS ~ nonverbalIQ
```{r nonverbal}

#using same priors

# fit model
model_n <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + b*nonverbalIQ_s ,
    a ~ dnorm( 15 , 2 ) ,
    b ~ dnorm( -0.5 , 0.5 ) ,
    sigma ~ dunif( 0 , 50 )
  ) ,  data=one )


#output with variance-covariance matrix
precis( model_n , corr=TRUE ) 

#plot
nonverbalIQ_seq <- seq( from= min(one$nonverbalIQ_s) , to=max(one$nonverbalIQ_s) , length.out=1000 ) 
pred=data.frame(nonverbalIQ_s=nonverbalIQ_seq)

mu <- link( model_n , n=1e4, data= pred)
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 ) 
sim_ADOS <- sim( model_n , data=pred ) 
ADOS_PI <- apply( sim_ADOS , 2 , PI , prob=0.89 ) 

#run all the lines together
plot( ADOS ~ nonverbalIQ_s , one , col=col.alpha(rangi2,0.5) ) 
lines( nonverbalIQ_seq , mu.mean ) 
shade( mu.PI , nonverbalIQ_seq ) 
shade( ADOS_PI , nonverbalIQ_seq ) 
```

model:
ADOS ~ socialIQ
```{r social}

#using same priors


# fit model
model_s <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + b*socialIQ_s ,
    a ~ dnorm( 15 , 2 ) ,
    b ~ dnorm( -0.5 , 0.5 ) ,
    sigma ~ dunif( 0 , 50 )
  ) ,  data=one )


#output with variance-covariance matrix
precis( model_s , corr=TRUE ) 

#plot
socialIQ_seq <- seq( from= min(one$socialIQ_s) , to=max(one$socialIQ_s) , length.out=1000 ) 
pred=data.frame(socialIQ_s=socialIQ_seq)

mu <- link( model_s , n=1e4, data= pred)
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 ) 
sim_ADOS <- sim( model_s , data=pred ) 
ADOS_PI <- apply( sim_ADOS , 2 , PI , prob=0.89 ) 

#run all the lines together
plot( ADOS ~ socialIQ_s , one , col=col.alpha(rangi2,0.5) ) 
lines( socialIQ_seq , mu.mean ) 
shade( mu.PI , socialIQ_seq ) 
shade( ADOS_PI , socialIQ_seq ) 
```

# Question 2
2. Do the different aspects of IQ account for different portions of the variance in ADOS? 
(knowing social IQ how much does it explane of AODS? and if you have 2?)

2.1. Does it make sense to have all IQ measures in the same model? First write a few lines answering the question and motivating your answer, including a discussion as to what happens when you put all of them in the same model. Then build a model following your answer. If your answer is "no", you are not free, you still have to answer: are there alternative ways of answering the question?

```{r corr}

#Looking at correlation
#verbal ~ nonverbal
model_vn <- map(
  alist(
    verbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bn*nonverbalIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bn ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=one )


precis(model_vn) 

# verbal ~ social
model_vs <- map(
  alist(
    verbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bs*socialIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bs ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=one )



precis(model_vs) 


# nonverbal - social
model_ns <- map(
  alist(
    nonverbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bs*socialIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bs ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=one )

precis(model_ns) 


#--
iq=data.frame(one$verbalIQ_s, one$nonverbalIQ_s, one$socialIQ_s)

plot(iq)

```


2.2. Build the model, assess its quality, write a few lines interpreting the results.

```{r model all}


model_all <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + bv*verbalIQ_s + bn*nonverbalIQ_s + bs*socialIQ_s,
    a ~ dnorm( 15 , 2 ) ,
    bv ~ dnorm( -0.5 , 0.5 ) ,
    bn ~ dnorm( -0.5 , 0.5 ) ,
    bs ~ dnorm( -0.5 , 0.5 ) ,
    sigma ~ dunif( 0 , 50 )
  ) ,  data=one )


#output with variance-covariance matrix
precis(model_all) 

plot(precis(model_all))

```



```{r}
samples_all = extract.samples( model_all , n=1e4 )

prediction <- sim( model_all , data=one, n= 1e4, post=samples_all )

library(bayesplot)
ppc_dens_overlay(one$ADOS, prediction[1:150, ])
#150 probable predictive posteriors
```


# Question 3

3. Let's now include also the TD children. Does it make sense to ask whether IQ and ADOS are related? Motivate your answer. In any case, if you wanted to build a model to answer that question, which model would you build? Run the model, assess its quality, write a few lines interpreting the results.

```{r clean and scale}

library(ggplot2)

ggplot(data = port) +
  geom_line(aes(x= ADOS), stat= "density") +
  facet_wrap(~ ASD)
  
#remove weird datapoints
sum(is.na(port))
h=na.omit(port)
sum(is.na(h))

i=subset(h, h$ASD == 0 & h$ADOS < 6  )
j=subset(h, h$ASD == 1 & h$ADOS > 6  )

cleaned= rbind(i, j)

#standardize (mean of 0, sd of 1)
cleaned$nonverbalIQ_s=(cleaned$nonverbalIQ - mean(cleaned$nonverbalIQ)) / sd(cleaned$nonverbalIQ)
cleaned$verbalIQ_s= (cleaned$verbalIQ - mean(cleaned$verbalIQ)) / sd(cleaned$verbalIQ)
cleaned$socialIQ_s= (cleaned$socialIQ - mean(cleaned$socialIQ)) / sd(cleaned$socialIQ)

ggplot(data = cleaned) +
  geom_line(aes(x= ADOS), stat= "density") +
  facet_wrap(~ ASD)
```

 
```{r plot}

#look at data to help understand it

#nonverbalIQ
ggplot(data = cleaned) +
  geom_point(aes(x= ADOS, y= nonverbalIQ_s)) +
  facet_wrap(~ ASD)

#verbalIQ
ggplot(data = cleaned) +
  geom_point(aes(x= ADOS, y= verbalIQ_s)) +
  facet_wrap(~ ASD)


#socialIQ
ggplot(data = cleaned) +
  geom_point(aes(x= ADOS, y= socialIQ_s)) +
  facet_wrap(~ ASD)



```

```{r}

model_td <- map(
  alist(
    ADOS ~ dnorm( mu , sigma ) ,
    mu <- a + bv*verbalIQ_s + bn*nonverbalIQ_s + bs*socialIQ_s + bd*ASD,
    a ~ dnorm( 10 , 4 ) ,
    bv ~ dnorm( 0 , 1 ) ,
    bn ~ dnorm( 0 , 1) ,
    bs ~ dnorm( 0 , 1 ) ,
    bd ~ dnorm( 0 , 5 ) , #binary variable - outcome will be the difference between TD and ASD
    sigma ~ dunif( 0 , 50 )
  ) ,  data=cleaned )


#output with variance-covariance matrix
precis(model_td) 

plot(precis(model_td))

```

```{r}
samples_td = extract.samples( model_td , n=1e4 )

prediction <- sim( model_td , data=cleaned, n= 1e4, post=samples_td )

library(bayesplot)
ppc_dens_overlay(cleaned$ADOS, prediction[1:150, ])
#150 probable predictive posteriors
```



# Question 4

4. Let's discuss contents:
4.1. You have three scores for IQ, do they show shared variance? Is that the same in TD and ASD? What does that tell us about IQ? (2 POV: 1) all IQ connected 2) you can be good at one and bad at other)
4.2. You have explored the relation between IQ and ADOS. How do you explain that relation from a cognitive perspective? N.B. You can present alternative hypotheses.

```{r corr}

#using visit 1 and children with ASD
asd= port[port$ASD== 1,]
#standardize (mean of 0, sd of 1)
asd$nonverbalIQ_s=scale(asd$nonverbalIQ)
asd$verbalIQ_s= scale(asd$verbalIQ)
asd$socialIQ_s= scale(asd$socialIQ)


#Looking at correlation
#verbal ~ nonverbal
model_vn <- map(
  alist(
    verbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bn*nonverbalIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bn ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=asd )


precis(model_vn) 

# verbal ~ social
model_vs <- map(
  alist(
    verbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bs*socialIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bs ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=asd )



precis(model_vs) 


# nonverbal - social
model_ns <- map(
  alist(
    nonverbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bs*socialIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bs ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=asd )

precis(model_ns) 


#--
# iq=data.frame(asd$verbalIQ_s, asd$nonverbalIQ_s, asd$socialIQ_s)
# 
# plot(iq)

```

```{r}
#using visit 1 and children with TD
td= port[port$ASD== 0,]
#standardize (mean of 0, sd of 1)
td$nonverbalIQ_s=scale(td$nonverbalIQ)
td$verbalIQ_s= scale(td$verbalIQ)
td$socialIQ_s= scale(td$socialIQ)

td=na.omit(td)

#Looking at correlation
#verbal ~ nonverbal
model_vn <- map(
  alist(
    verbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bn*nonverbalIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bn ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=td )


precis(model_vn) 

# verbal ~ social
model_vs <- map(
  alist(
    verbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bs*socialIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bs ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=td )



precis(model_vs) 


# nonverbal - social
model_ns <- map(
  alist(
    nonverbalIQ_s ~ dnorm( mu , sigma ) ,
    mu <- a + bs*socialIQ_s,
    a ~ dnorm( 0 , 1 ) ,
    bs ~ dnorm( 0 , 1 ) ,
    sigma ~ dunif( 0 , 1 )
  ) ,  data=td )

precis(model_ns) 



```


5. Bonus questions: Including measurement errors. 
5.1. Let's assume that ADOS has a measurement error of 1. How would you model that in any of the previous models? 
5.2. We know that IQ has an estimated measurement error of 2.12. How would you include that? 


